\bf Appendix: Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-04-20 appendix.qmd
Added dobious cells

For illustration purposes, we only selected cells generated with inDrops ($n=5858$ cells) and UMAP cell embeddings. Also, @xia2023 used first $9$ principal components to generate the UMAP. The objective is to assess the optimized layout by scDEED, and if it does not accurately represent the three clusters with small separations of the PBMC dataset, then select a reasonable \gD{} layout.

The layout a (@fig-pbmc-mse-umap) is generated from the hyper-parameters suggested by @chen2023, and the layout b (@fig-pbmc-mse-umap) is with suggested hyper-parameters by scDEED to be more accurate. The layout a and b contain $46$ and $15$ dubious cells respectively.

@@

```{r}
#| label: fig-pbmc-mse-umap
#| fig-pos: H
#| out-width: 100%
#| fig-cap: "Assessing which of the two UMAP layouts with different hyper-parameter settings (n_neighbors: $30$, min_dist: $0.3$ (red); n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the top plot and in the scatterplots represents UMAP layout (a, b). Of the two, layout b is optimal across all binwidths making it the best choice."

>>>>

For illustration purposes, we only selected cells generated with inDrops ($n=5858$ cells) and UMAP and tSNE cell embeddings. Also, @xia2023 used first $9$ principal components to generate the UMAP and tSNE. The objective is to assess the optimized layout by scDEED, and if it does not accurately represent the three clusters with small separations of the PBMC dataset, then select a reasonable \gD{} layout.

The layout a (@fig-pbmc-mse-umap) is generated from the hyper-parameters suggested by @chen2023, and the layout b (@fig-pbmc-mse-umap) is with suggested hyper-parameters by scDEED to be more accurate. The layout a and b contain $46$ and $83$ dubious cells respectively.

@@

```{r}
#| label: fig-pbmc-mse-umap
#| fig-pos: H
#| out-width: 100%
#| fig-cap: "Assessing which of the two layouts with UMAP and tSNE with different hyper-parameter setting (n_neighbors: $30$, min_dist: $0.3$ (red); n_neighbors: $30$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the top plot and in the scatterplots represents UMAP and tSNE layouts (a, b). Of the two, layout b is optimal across all binwidths making it the best choice."

### 2025-04-20 appendix.tex
Added dobious cells

For illustration purposes, we only selected cells generated with inDrops
(\(n=5858\) cells) and UMAP cell embeddings. Also, \citet{xia2023} used
first \(9\) principal components to generate the UMAP. The objective is
to assess the optimized layout by scDEED, and if it does not accurately
represent the three clusters with small separations of the PBMC dataset,
then select a reasonable \gD{} layout.

The layout a (Figure~\ref{fig-pbmc-mse-umap}) is generated from the
hyper-parameters suggested by \citet{chen2023}, and the layout b
(Figure~\ref{fig-pbmc-mse-umap}) is with suggested hyper-parameters by
scDEED to be more accurate. The layout a and b contain \(46\) and \(15\)
dubious cells respectively.

@@

\caption{\label{fig-pbmc-mse-umap}Assessing which of the two UMAP
layouts with different hyper-parameter settings (n\_neighbors: \(30\),
min\_dist: \(0.3\) (red); n\_neighbors: \(80\), min\_dist: \(0.5\)
(blue)) on the PBMC data is the better representation using MSE for
varying binwidth (\(a_1\)). Colour used for the lines and points in the
top plot and in the scatterplots represents UMAP layout (a, b). Of the
two, layout b is optimal across all binwidths making it the best
choice.}

>>>>

For illustration purposes, we only selected cells generated with inDrops
(\(n=5858\) cells) and UMAP and tSNE cell embeddings. Also,
\citet{xia2023} used first \(9\) principal components to generate the
UMAP and tSNE. The objective is to assess the optimized layout by
scDEED, and if it does not accurately represent the three clusters with
small separations of the PBMC dataset, then select a reasonable \gD{}
layout.

The layout a (Figure~\ref{fig-pbmc-mse-umap}) is generated from the
hyper-parameters suggested by \citet{chen2023}, and the layout b
(Figure~\ref{fig-pbmc-mse-umap}) is with suggested hyper-parameters by
scDEED to be more accurate. The layout a and b contain \(46\) and \(83\)
dubious cells respectively.

@@

\caption{\label{fig-pbmc-mse-umap}Assessing which of the two layouts
with UMAP and tSNE with different hyper-parameter setting (n\_neighbors:
\(30\), min\_dist: \(0.3\) (red); n\_neighbors: \(30\) (blue)) on the
PBMC data is the better representation using MSE for varying binwidth
(\(a_1\)). Colour used for the lines and points in the top plot and in
the scatterplots represents UMAP and tSNE layouts (a, b). Of the two,
layout b is optimal across all binwidths making it the best choice.}
